---
title: "volcano3D package"
author: "Katriona Goldmann"
output: md_document
vignette: >
  %\VignetteIndexEntry{volcano3D package}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.height = 7, 
                      fig.width=7, 
                      fig.align = "center")
library(knitr)
library(kableExtra)
```


[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/volcano3D)](https://cran.r-project.org/package=volcano3D)
[![Codecov test coverage](https://codecov.io/gh/r-lib/volcano3D/branch/master/graph/badge.svg)](https://codecov.io/gh/r-lib/volcano3D?branch=master)
[![Downloads](https://cranlogs.r-pkg.org/badges/volcano3D?color=blue)](https://cran.rstudio.com/package=volcano3D)

# Introduction

```{r, echo=FALSE}
library(ggplot2)
library(ggpubr)
library(plotly)
library(usethis)
```


The volcano3D package enables exploration of probes differentially 
expressed between three groups. Its main purpose is for the
visualisation of differentially expressed genes in a three-dimensional
volcano plot. These plots can be converted to interactive visualisations using
plotly. 

The vignette explores a case study from the PEAC 
rheumatoid arthritis trial (Pathobiology of Early Arthritis Cohort). 
The methodology has been published in 
['Lewis, Myles J., et al. "Molecular portraits of early rheumatoid arthritis 
identify clinical and treatment response phenotypes." Cell reports 28.9 (2019): 
2455-2470.'](https://www.cell.com/cell-reports/fulltext/S2211-1247(19)31007-1) 
with an interactive web tool available at 
[https://peac.hpc.qmul.ac.uk](https://peac.hpc.qmul.ac.uk).  

This tool acts as a searchable interface to examine relationships between individual 
synovial and blood gene transcript levels and histological, clinical, and 
radiographic parameters, and clinical response at 6 months. 
An interactive interface allows the gene module analysis to be explored for 
relationships between modules and clinical parameters. 
The PEAC interactive web tool was creating as an 
[R Shiny app](https://shiny.rstudio.com) and deployed to the web using a server. 


## Installing and Loading the package

#### Prerequisite

* ggplot2 
* ggpubr 
* ggrepel
* plotly

### Install from CRAN

*Not yet publicly available:*

```{r, eval = FALSE}
install.packages("volcano3D")
```


### Install from Github

*Not yet publicly available:*

```{r, eval = FALSE}
library(devtools)
install_github("KatrionaGoldmann/volcano3D")
```

---

# Citation

volcano3D was developed by the bioinforamtics team at the 
[Experimental Medicine & Rheumatology department](https://www.qmul.ac.uk/whri/emr/) 
and [Centre for Translational Bioinformatics](https://www.qmul.ac.uk/c4tb/) at 
Queen Mary University London. 

If you use this package please cite as: 

> Lewis, Myles J., et al. "Molecular portraits of early rheumatoid arthritis 
identify clinical and treatment response phenotypes." Cell reports 28.9 (2019): 
2455-2470.

or using:

```{r}
citation("volcano3D")
```

